---
title: "Homework 1 - Time Series Visualization"
author: "Prof. Dr. Stephan Trahasch"
date: 'Submission date: 17.12.2020'
output:
  html_document:
    theme: cerulean
    css: styles/homework.css
    toc: yes
    toc_depth: 3
    toc_float: yes
  word_document:
    toc: yes
    toc_depth: '3'
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
options(digits = 5)

# libraries to load
library(fpp3)
library(fpp2) 
```

# Exercise 1
Use the help menu to explore what the series `gold`, `woolyrnq` and `gas` represent. These are available in the `forecast` package.

```{r, echo=TRUE, cache=TRUE, eval=FALSE}
help(gold)  # Daily morning gold prices in US dollars. 1 January 1985 - 31 March 1989.
# to do
help(woolyrnq)  # Quarterly production of woollen yarn in Australia: tonnes. Mar 1965 - Sep 1994.
help(gas)  # Australian monthly gas production: 1956-1995.
```

Use `autoplot` to plot each of these in separate plots.

```{r, echo=TRUE, cache=TRUE}
# to do
require(gridExtra)
plot1 <- autoplot(gold)
plot2 <- autoplot(woolyrnq)
plot3 <- autoplot(gas)
grid.arrange(plot1, plot2, plot3)

```

**What is the frequency of each commodity series?**
Hint: apply the `frequency()` function.

```{r, echo=TRUE, cache=TRUE}
# to do
print("1  = Pro Tag 1 Wert. 4 = 4 Quartale pro Jah. 12 = 12 Monate pro Jahr")
print("gold: ")
frequency(gold)
print("woolyrnq: ")
frequency(woolyrnq)
print("Gas: ")
frequency(gas)
```

Use `which.max()` to spot the outlier in the `gold` series. 
**Which observation was it?**

```{r, echo=TRUE, cache=TRUE}
# to do
print("Am Tag: ")
which.max(gold)
print("Betrag: ")
gold[770]
```

# Exercise 2 

Import the file `tute1.csv` from the folder data and review its contents. You should find four columns of information. Columns B through D each contain a quarterly series, labelled Sales, AdBudget and GDP. Sales contains the quarterly sales for a small company over the period 1981-2005. AdBudget is the advertising budget and GDP is the gross domestic product. All series have been adjusted for inflation.

You can read the data into R with the following script:

```{r, echo=TRUE, cache=TRUE}
tute1 <- read.csv("data/tute1.csv", header = TRUE)
View(tute1)
```

We convert the data to time series. The `[,-1]` removes the first column which contains the quarters as we don't need them now.

```{r, echo=TRUE, cache=TRUE}
mytimeseries <- ts(tute1[, -1], start = 1981, frequency = 4)
View(mytimeseries)
```

Construct time series plots of each of the three series and use 'facets = TRUE' .

```{r, echo=TRUE, cache=TRUE}
# to do
autoplot(mytimeseries, facets = TRUE)
```

Check what happens when you don't include `facets=TRUE`.

```{r, echo=TRUE, cache=TRUE, dependson="exmyts"}
# to do
autoplot(mytimeseries)
print("Facets teil es in unterschiedliche Grids ein")
```

# Exercise 3

Import some monthly Australian retail data from the folder data. These represent retail sales in various categories for different Australian states, and are stored in a MS-Excel file.

You can read the data into R with the following script:

```{r, echo=TRUE, cache=TRUE}
retaildata <- readxl::read_excel("data/retail.xlsx", skip = 1)
View(retaildata)
```

The second argument (`skip=1`) is required because the Excel sheet has two header rows.

Select one of the time series as follows (but replace the column name with your own chosen column):

```{r}
colnames(retaildata)
```


```{r, echo=TRUE, cache=TRUE}
myts <- ts(retaildata[, "A3349432V"], frequency = 12, start = c(1982, 4))
```

Explore your chosen retail time series using the following functions: `autoplot`, `ggseasonplot`, `ggsubseriesplot`, `gglagplot`, `ggAcf`

```{r, echo=TRUE, cache=TRUE}
# to do
autoplot(myts)
ggseasonplot(myts)
ggsubseriesplot(myts)
gglagplot(myts)
ggAcf(myts)
```

**Can you spot any seasonality, cyclicity and trend? What do you learn about the series?**

Your answer:
- Sales sind steigend (trend nach oben)
- Sales nehmen zum Ende des Jahres zu (seasonality)
- Daten haben eine starke Korrelation
- cyclicity konnten keine gefunden werden

(Weihnachtsshopping)

# Exercise 4
Create time plots of the following time series: goog, writing, fancy, a10, h02.

    + Use `help()` to find out about the data in each series.
    + For the `goog` plot, modify the axis labels and title.

```{r, echo=TRUE, message=FALSE, warning=FALSE}
# to do
help(goog) # Daily closing stock prices of Google Inc
help(writing) #Sales of printing and writing paper
help(fancy) # Monthly sales for a souvenir shop on the wharf at a beach resort town in Queensland, Australia
help(a10) # Monthly anti-diabetic drug subsidy in Australia from 1991 to 2008.
help(h02) # Monthly corticosteroid drug subsidy in Australia from 1991 to 2008

autoplot(goog) + labs(x = "Day", y="Price in USD", title= "Daily closing stock prices of Google Inc")
autoplot(writing)
autoplot(fancy)
autoplot(a10)
autoplot(h02)
```

Use the `ggseasonplot` and `ggsubseriesplot` functions to explore the seasonal patterns in the following time series:  `writing`, `fancy`, `a10`, `h02`.

- What can you say about the seasonal patterns?
- Can you identify any unusual years?


```{r, echo=TRUE}
# writing
# to do
ggseasonplot(writing)
ggsubseriesplot(writing)
```

Your answer:
Deutlicher sessionaler Merkmal zu erkennen. Abstieg im August. wahrscheinlich wegen Sommerferien

```{r, echo=TRUE}
# fancy
# to do
ggseasonplot(fancy)
ggsubseriesplot(fancy)
```

Your answer:#
Deutlich mehr Sale im november, Dezember (Weihnachtszeit/Urlaub). Es handelt sich um ein Souvenirshop

```{r, echo=TRUE}
# a10
# to do
ggseasonplot(a10)
ggsubseriesplot(a10)
help(a10)
```

Your answer:
Deutlicher Anstieg nach der Weihnachtszeit. Evlt. Insulin Spritze.
Bei 2008 hört die Datenreihe auf.

```{r, echo=TRUE}
# h02
# to do
ggseasonplot(h02)
ggsubseriesplot(h02)
help(h02)
```

Your answer:
Febraur - Januar stark steigend, dann stark fallend, da

# Exercise 5
Use the following graphics functions: `autoplot`, `ggseasonplot`, `ggsubseriesplot`, `gglagplot`, `ggAcf` and explore features from the following time series: `hsales`, `usdeaths`, `bricksq`, `sunspotarea`, `gasoline`.

    + Can you spot any seasonality, cyclicity and trend?
    + What do you learn about the series?

Example for hsales:

```{r, echo=TRUE, cache=TRUE}
autoplot(hsales)
ggseasonplot(hsales)
ggsubseriesplot(hsales)
gglagplot(hsales)
ggAcf(hsales)
```

  + Seasonality evident in all plots
  + Cyclicity seen in first two plots
  + No trend
  + ACF only shows seasonality. Cycle length too long to show up here.


```{r, echo=TRUE, cache=TRUE}
# to do
autoplot(usdeaths)
ggseasonplot(usdeaths)
ggsubseriesplot(usdeaths)
gglagplot(usdeaths)
ggAcf(usdeaths)
```
Your answer:
 + Seasonality in summer
 + No Trend
 + No Cyclicity
 + ACF only shows seasonality
 
 + Probably more accidental deaths in summer, because people are more outside. (Drive more, swimming and so on)


```{r, echo=TRUE, cache=TRUE}
# to do
autoplot(bricksq)
ggseasonplot(bricksq)
ggsubseriesplot(bricksq)
gglagplot(bricksq)
ggAcf(bricksq)

```

Your answer:
  + Small Seasonality (lower in Q1)
  + Cyclicity from 1975 good shown in ggsubseriesplot
  
  + AFC shows seasonality but no cyclicity
  
```{r}
 View(sunspotarea)
```

  
```{r, echo=TRUE, cache=TRUE}
# to do
autoplot(sunspotarea)
gglagplot(sunspotarea, lag=70)
ggAcf(sunspotarea)

```

Your answer:
  + Cyclicity ~ around every 10 years it drops
  + no seasonality
  + no trend
  
  + Every few years some physical stuff is happening with the sun/earth




```{r, echo=TRUE, cache=TRUE}
# to do
autoplot(gasoline)
ggseasonplot(gasoline)
gglagplot(gasoline)
ggAcf(gasoline, lag.max = 200)


```

Your answer:
  + seasonality every half year 
  + maybe cyclicity with falling graph after 2009 

# Exercise 6

The `pigs` data shows the monthly total number of pigs slaughtered in Victoria, Australia, from Jan 1980 to Aug 1995. Use `mypigs <- window(pigs, start=1990)` to select the data starting from 1990. Use `autoplot` and `ggAcf` for `mypigs` series and compare these to white noise plots.

```{r, echo=TRUE, cache=TRUE}
# White noise
set.seed(1711)
y <- ts(rnorm(50))
autoplot(y) + ggtitle("White noise")
ggAcf(y)

```



```{r, echo=TRUE, cache=TRUE}
# todo
mypigs <- window(pigs, start=1990)
autoplot(mypigs)
ggAcf(mypigs)
```

Your answer:
With a higher starting year the ACF plot goes above the blue dotted index line, confirming that the data has no white noise(see lag 1,2,3)

maybe because of lagg 12 there could be some seasonality too, but only a slight one.

# Exercise 7

`dj` contains 292 consecutive trading days of the Dow Jones Index. Use `ddj <- diff(dj)` to compute the daily changes in the index. Plot `ddj` and its ACF. Do the changes in the Dow Jones Index look like white noise?




```{r}
# todo
ddj <- diff(dj)
autoplot(ddj)
ggAcf(ddj, lag=200)
```

Your answer:

The Acf shows that there should be no white noise, because it crossea the blue limit line sometimes. But in the first diagram its kinda hard to see it. 
Probably if you add more data, the white noise should get filtered out even more.

